AA Distributed Software Development Methodology

### 2013-06-02 paper.tex
moving on to conclusions

Easy and effective GSD team management is the main
purpose of the AA methodology. We applied the proposed methodology to a group
of 9 developers in July of 2011 on what we called Lab
Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}} The
main objective of the team was to work on an array of strategic free software
projects in the broad audiovisual and web categories,
contributing directly to their development, submitting bug patches or
committing new features to their source code.

The team members had different levels of experience on software
development for large and distributed free software projects
like Scilab or Mozilla. In this way, one month of training was
conducted by two experienced developers, teaching the basics of use of
development support tools like bugtrackers, programming languages,
version control systems, and build systems. After this period, a starter project
was proposed for new developers, namely to have an accepted patch or commit to
the official repository for a large and widely used free software app.
Developers passing the starter project would be deemed `initiated' and be
called a `Macambira' developer.  Table~\ref{tabela:contribuicoes} summarizes the
effective accepted contributions of each successfully initiated developer to
free software projects in 2011, which used the AA methodology.

@@

\begin{table}
  \caption{Free and open software projects that received contributions
    from successfully initiated developers  or `macambiras' that worked under
    the AA methodology. The first column lists applications to which
    contributions were officially accepted and whose development process was
    performed and logged using AA. The second column shows the pseudonym of the
    ``Macambiras'' who sent \textit{commits} to the application. At
    Lab Macambira, and at the free software community at large, 
    It is
    common practice at Lab Macambira, and the free software community at large,
    to use pseudonyms as identification in order to enhance privacy in practice.}
    \begin{tabular}{|l|l|}
        \hline
        Application           & ``Commiters''                       \\
        \hline \hline
        Mozilla Firefox       & daneoshiga, bzum                    \\
        Evince                & hick209, bzum, marcicano, mquasar   \\
        BePDF / Xpdf          & marcicano                           \\
        Ekiga                 & flecha                              \\
        Empathy               & fefo                                \\
        Lib Folks (Telepathy) & kamiarc                             \\
        Scilab                & v1z, humannoise                     \\
        VxL                   & v1z                                 \\
        ImageMagick           & v1z                                 \\
        OpenOffice            & hick209                             \\
        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
        Puredata OpenCV       & v1z                                 \\
        Puredata GEM          & v1z, fefo, hick209                  \\
        Puredata PDP          & v1z, fefo, hick209                  \\
        ChucK                 & rfabbri, automata                   \\
        ChucK MiniAudicle     & rfabbri, automata                   \\
        WebRTC                & automata                            \\
        OSC-Web               & automata                            \\
        Web-PD-GUI            & automata                            \\
        Live-Processing       & automata                            \\
        ChucK-Wiimote         & automata                            \\
        Audiolet              & automata                            \\
        Extempore             & automata                            \\
        \hline
    \end{tabular}
    \label{tabela:contribuicoes}
\end{table}

In one month, each developer officially contributed to one or many free software
projects. Many developers started the initiation training with no knowledge of what
is free software and ended that period becoming a free software developer.
During that month, the same team of trainees also developed the first version of the AA
system and used AA to manage their activities, even while developing
the other aforementioned free software projects.

The source code of AA --- both the client that sends the
logs and the AA Web server --- is public available as free software\footnote{AA source
  code:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
and all the actual AA log data of the entire team of Lab Macambira from 2011 to
the present time is also on-line\footnote{Logs of AA sessions:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.

After the initial training period of 1 month, the ``Macambiras''
worked during 6 additional months on a large range of free software projects, distributed on work
groups --- each work group focusing on a specific theme like video,
audio and web --- and funded by contracts, freelance, and support of the
Pont\~{a}o N\'{o}s Digitais\footnote{\url{http://nosdigitais.teia.org.br}}. 
Table~\ref{tabela:criados} has a list of \emph{new} free software applications created by
``Lab Macambira'' since July 2011.

\begin{table}
    \caption{Software projects created by Lab Macambira since July
      2011 with a short description and the technologies involved (\eg,
      programming languages or frameworks). It is
      interesting to note the heterogeneity of projects and their areas
      of application.}

    \footnotesize{
    \begin{tabular}{|l|p{5cm}|l|}
        \hline
        Application & Description & Technologies involved \\ 
        \hline \hline
        AA            & Algorithmic Auto-regulation      & Python, PHP \\
        \hline
        \'{A}gora Communs & System for on-line deliberations & PHP \\
        \hline
        SIP           & Scilab Image Processing toolbox & C, Scilab \\
        \hline
        animal        & An Imaging Library              & C \\
        \hline
        TeDi          & Test Framework for Distance Transform
        Algorithms & C, Shellscript, Scilab \\
        \hline
        Macambot      & Multi-use IRC Bot               & Python \\
        \hline
        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
        conference of the rights of minors & PHP, JavaScript \\
        \hline
        CPC           & Center for accounting of the Brazilian culture
        representation groups & Python, Django \\
        \hline
        Timeline      & Interactive time lines on the Web & JavaScript
        \\
        \hline
        Imagemap      & Interactive marking for on-line photos &
        JavaScript \\
        \hline
        ABT           & Program for real-time execution and musical
        rhythmic analysis & Python \\
        \hline
        EKP           & Emotional Kernel Panic & Python, ChucK \\
        \hline
        SOS           & Aggregation and diffusion of popular and native
        knowledge about health & Python, Django \\
        \hline
        Creative Economy & Platform for creative, collaborative and
        solidarity economy of the culture hubs and cultural entities &
        Python, Django \\
        \hline
        OpenID Integration & Adaptations to existing software for
        unified login through OpenID & PHP \\
        \hline
        pAAinel & Dashboard for the real-time visualization of Lab
        Macambira activity & Python, Django \\
        \hline
        Georef & Collection of scripts to be used as reference, which
        aims to be a GIS platform to map public data of use to
        citizens & Python, Django \\
        \hline
        AirHackTable & Software for an instrument which generates
        sound from flying origami tracked by webcams & Puredata,
        C/C++, Scilab \\
        \hline
        \end{tabular}}
    \label{tabela:criados}
\end{table}

As of this writing the `Lab Macambira' unites 15 software developers, and 
key 2011-initiated developers continue to work voluntarily in the project.

>>>>

Easy and effective GSD team management is the main purpose of the AA
methodology. We applied the proposed methodology to a group of 9 developers in
July of 2011, 3 of which are coauthors of the present work, on what we called Lab
Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}} The
main objective of the team was to work on an array of strategic free software
projects in the broad audiovisual and web categories, contributing directly to
their development, submitting bug patches or committing new features to their
source code.

The team members had different levels of experience on software development for
large and distributed free software projects like Scilab or Mozilla. In this
way, one month of training was conducted by three experienced developers (the
first authors of the present work), teaching
the basics of use of development infrastructure tools like bugtrackers, programming
languages, version control systems, and build systems. After this period, a
starter project was proposed for new developers: to submit a bugfix or implement
a new feature and have an accepted
patch or commit to the official repository for a large free
software project.  Developers passing the starter project would be deemed
`initiated' and be called a `Macambira' developer, and be hired for the
remainder of the semester.
Table~\ref{tabela:contribuicoes} summarizes the effective accepted contributions
of each successfully initiated developer to free software projects in 2011,
which used the AA methodology.

@@

\begin{table}
  \caption{Free and open software projects that received contributions from
  successfully initiated developers  or `Macambiras' that worked under the AA
  methodology. The first column lists applications to which contributions were
  officially accepted and whose development process was tracked using AA. The
  second column shows the pseudonym of the committers. At Lab Macambira it is
  common practice to use pseudonyms in AA as identification in order to enhance
  privacy.}
  \begin{center}
  \small\begin{tabular}{|l|l|}
        \hline
        Application           & ``Commiters''                       \\
        \hline \hline
        Mozilla Firefox       & daneoshiga, bzum                    \\
        Evince                & hick209, bzum, marcicano, mquasar   \\
        BePDF / Xpdf          & marcicano                           \\
        Ekiga                 & flecha                              \\
        Empathy               & fefo                                \\
        Lib Folks (Telepathy) & kamiarc                             \\
        Scilab                & v1z, humannoise                     \\
        VxL                   & v1z                                 \\
        ImageMagick           & v1z                                 \\
        OpenOffice            & hick209                             \\
        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
        Puredata OpenCV       & v1z                                 \\
        Puredata GEM          & v1z, fefo, hick209                  \\
        Puredata PDP          & v1z, fefo, hick209                  \\
        ChucK                 & rfabbri, automata                   \\
        ChucK MiniAudicle     & rfabbri, automata                   \\
        WebRTC                & automata                            \\
        OSC-Web               & automata                            \\
        Web-PD-GUI            & automata                            \\
        Live-Processing       & automata                            \\
        ChucK-Wiimote         & automata                            \\
        Audiolet              & automata                            \\
        Extempore             & automata                            \\
        \hline
    \end{tabular}
    \end{center}
    \label{tabela:contribuicoes}
\end{table}

In one month, each developer officially contributed to one or many free software
projects. Many developers started the initiation training with no knowledge of what
is free software and ended that period becoming a free software developer.
During that month, the same team of trainees also developed the first version of the AA
system and used AA to manage their activities, even while developing
the other aforementioned free software projects. Thus, AA and the associated
software system was tested, prototyped, and developed in close contact with
actual practice.  The source code of AA --- both the client that sends the
logs and the AA Web server --- is public available as free software\footnote{AA source
  code:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
and all the actual AA log data of the entire team of Lab Macambira from 2011 to
the present time is also on-line\footnote{Logs of AA sessions:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.

After the initial training period of 1 month, the initiated ``Macambiras''
worked during 6 additional months on a large range of free software projects,
divided into work groups --- each work group focusing on a specific theme like
video, audio and web --- and funded by contracts, freelance, and support of the
Pont\~{a}o N\'{o}s Digitais\footnote{\url{http://nosdigitais.teia.org.br}}.
Table~\ref{tabela:criados} has a list of \emph{new} free software applications
created by ``Lab Macambira'' since July 2011.

\begin{table}
    \caption{Software projects created by Lab Macambira since July
      2011 with a short description and the technologies involved (\eg,
      programming languages or frameworks). It is
      interesting to note the heterogeneity of projects and their areas
      of application.}
    \footnotesize{\begin{tabular}{|l|p{5cm}|l|}
        \hline
        Application & Description & Technologies involved \\ 
        \hline \hline
        AA            & Algorithmic Auto-regulation      & Python, PHP \\
        \hline
        \'{A}gora Communs & System for on-line deliberations & PHP \\
        \hline
        SIP           & Scilab Image Processing toolbox & C, Scilab \\
        \hline
        animal        & An Imaging Library              & C \\
        \hline
        TeDi          & Test Framework for Distance Transform
        Algorithms & C, Shellscript, Scilab \\
        \hline
        Macambot      & Multi-use IRC Bot               & Python \\
        \hline
        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
        conference of the rights of minors & PHP, JavaScript \\
        \hline
        CPC           & Center for accounting of the Brazilian culture
        representation groups & Python, Django \\
        \hline
        Timeline      & Interactive time lines on the Web & JavaScript
        \\
        \hline
        Imagemap      & Interactive marking for on-line photos &
        JavaScript \\
        \hline
        ABT           & Program for real-time execution and musical
        rhythmic analysis & Python \\
        \hline
        EKP           & Emotional Kernel Panic & Python, ChucK \\
        \hline
        SOS           & Aggregation and diffusion of popular and native
        knowledge about health & Python, Django \\
        \hline
        Creative Economy & Platform for creative, collaborative and
        solidarity economy of the culture hubs and cultural entities &
        Python, Django \\
        \hline
        OpenID Integration & Adaptations to existing software for
        unified login through OpenID & PHP \\
        \hline
        pAAinel & Dashboard for the real-time visualization of Lab
        Macambira activity & Python, Django \\
        \hline
        Georef & Collection of scripts to be used as reference, which
        aims to be a GIS platform to map public data of use to
        citizens & Python, Django \\
        \hline
        AirHackTable & Software for an instrument which generates
        sound from flying origami tracked by webcams & Puredata,
        C/C++, Scilab \\
        \hline
        \end{tabular}}
    \label{tabela:criados}
\end{table}

As of this writing, `Lab Macambira' unites over 15 software developers, and 
key developers trained in 2011 continue to work voluntarily in the project.
